Quantifying maternal reproductive output of chondrichthyan fishes

### 2024-05-19 README.qmd
readme
revisions

---
title: "Maternity Function"
format: gfm
editor: visual
---

## Quarto

@@

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

>>>>

---
title: "Quantifying maternal reproductive output of chondrichthyan fishes"
format: gfm
---

@@

## Example

Load data for sandbar shark maternity data set. Rows are data from individuals sharks x is fork length (cm) and z is maternal status (0 = non maternal, 1 = maternal).

```{r message=FALSE, warning=FALSE}
library(tidyverse)

data <- read_rds(here::here("data", "empirical-plumbeus.Rds"))
```

## Bin data into suitable length categories for plotting

```{r message=FALSE, warning=FALSE}
brks = seq(40, 220, 7.5)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())
```

Compile 3 parameter logistic model

```{r message=FALSE, warning=FALSE}
library(TMB)
compile("code/logistic3.cpp")
dyn.load(dynlib("code/logistic3"))
```

### 2024-05-19 rmds/figures-supp.Rmd
revisions

data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples)))
```

@@

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "m50")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Precision - Length at 50% maternity")) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S4. Precision (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "c")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Precision - "~italic(P)[Max])) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S5. Precision (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets. Note 2PLF-maturity and 3PLF-fixed methods both used pre-specified values for $P_{Max}$ and are identical. 
\newpage

@@

```{r fig.width = 12, fig.height = 10}
  ggplot(data %>% filter(par_name %in% "r_0")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % relative absolute error", title = "Precision - Reproductive Output") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S11. Precision (per cent absolute error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets

>>>>

data <- read_rds(here("data", "simulations-summarised.Rds")) %>% 
  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples)))
```

@@

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "m50")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Accuracy - Length at 50% maternity")) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S4. Accuracy (per cent absolute error) in parameter estimates of $\hat{L'_{50}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets.

```{r, fig.width = 12, fig.height = 10}
ggplot(data %>% filter(par_name %in% "c")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % absolute relative error", title = expression("Accuracy - "~italic(P)[Max])) +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S5. Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ using alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets. Note 2PLF-maturity and 3PLF-fixed methods both used pre-specified values for $P_{Max}$ and are identical. 
\newpage

@@

```{r fig.width = 12, fig.height = 10}
  ggplot(data %>% filter(par_name %in% "r_0")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(method_long ~ species + mesh_name_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Mean % relative absolute error", title = "Accuracy - Reproductive Output") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure S11. Accuracy (per cent absolute error) in $R_{0}$ calculated using maternity parameters obtained from alternative logistic regression models. Each point reflects a mean value from 300 simulated data sets

### 2024-05-19 rmds/figures.Rmd
revisions

data %>% 
  filter(par_name %in% "c") %>% 
  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
  filter(mesh_name %in% "high") %>% 
  ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(species ~ method_long) + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  theme_bw() + 
  labs(x = "Sample size", y = "Mean % absolute relative error") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure 3. Precision (per cent absolute error) in parameter estimates  of $\hat{P_{Max}}$ for 3PLF methods. Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and precision decreased as the duration of the reproductive cycle increased. Each point reflects a mean value from 300 simulated data sets. 
\
\
\newpage

@@

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs(rel_error)) %>% 
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure 5. Performance of alternative maternity functions in minimising bias in calculations of $R_0$. The preferred method was that which minimised bias, $|\text{relative error}|$ across 300 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
\
\
\newpage

@@

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs_error) %>%
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure 6. Performance of alternative maternity functions in accurately calculating $R_0$. The best performing method was that which minimised mean absolute error across 300 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
\
\newpage

@@

# Bin data into length categories for plotting
brks = seq(40, 220, 7.5)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

@@

Figure 7. Comparison of 3PLF-estimated and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C. plumbeus*, in the Gulf of Mexico and Western North Atlantic. Solid line is the expected proportion in maternal condition at length, $\Psi'(L)$. The grey shaded region denotes 95% confidence intervals based on bootstrap resampling. $P_{Max}$ was fixed at 0.5 in the lower panel

>>>>

data %>% 
  filter(par_name %in% "c") %>% 
  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
  filter(mesh_name %in% "high") %>% 
  ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(species ~ method_long) + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  theme_bw() + 
  labs(x = "Sample size", y = "Mean % absolute relative error") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure 3. Accuracy (per cent absolute error) in parameter estimates  of $\hat{P_{Max}}$ for 3PLF methods. Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased. Each point reflects a mean value from 300 simulated data sets. 
\
\
\newpage

@@

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs(rel_error)) %>% 
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure 5. Performance of alternative maternity functions in minimising bias in calculations of $R_0$. The preferred method was that which minimised bias, $|\text{relative error}|$ across 300 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
\
\
\newpage

@@

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(factor(mesh_name, levels = c("low", "high")), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs_error) %>%
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure 6. Performance of alternative maternity functions in accurately calculating $R_0$. The best performing method was that which minimised mean absolute error across 300 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
\
\newpage

@@

# Bin data into length categories for plotting
brks = seq(40, 220, 10)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

@@

Figure 7. Comparison of 3PLF-estimated and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C. plumbeus*, in the Gulf of Mexico and Western North Atlantic. Solid line is the expected proportion in maternal condition at length, $\Psi'(L)$. The grey shaded region denotes 95% confidence intervals based on bootstrap resampling. Black points show proportion in maternal condition in 10cm length intervals. Marginal rug plots denote raw data that models were fit to. $P_{Max}$ was fixed at 0.5 in the lower panel.
